Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-07-02 ms.tex
Adding new plots for escape time distribution and frequency dependence

We now compare each of the solutions for surface flux with Monte Carlo simulations. In Fig. \ref{fig:sol_mc_residual_0}, the Monte Carlo is shown with solutions $H_{\rm d}$, $H_{\rm 0}$, and $H_{\rm 0 + bc} = H_{\rm 0} + H_{\rm bc}$ over-plotted for an optical depth of $\tau_0 = 10^7$ and photons emitted at line center $\rm x_s = 0$. The solution with the correct frequency-dependent boundary condition enforced, $H_{\rm 0 + bc}$, has considerably lower residuals to Monte Carlo than the other solutions, especially in the line wing. It is evident that the divergent solution $H_{\rm d}$ fails in the line wings. The $H_{\rm bc}$ term, when added to the solution $H_0$ derived by \citet{2006ApJ...649...14D}, adequately corrects for the apparent excess of flux in the peaks of the spectrum. This peak excess is also present in previous work that used the \citet{1990ApJ...350..216N} solution, such as \citet{2015MNRAS.449.4336S}. The boundary term resolves the slight deficit of $H_{\rm0}$ in the line wings, further improving agreement with the numerical result. Note that the residuals to the $H_0$ solution are a close match to the $H_{\rm bc}$ term, since the Monte Carlo represents the ``true'' solution, $H$, and $H_{\rm bc} = H - H_0$.

@@

@@

Integrating $dE/dt d\nu$ from Eq. \ref{eq:dEdtdnu} over all times should give a distribution for the emitted frequencies. It is expected that this time-integrated spectrum of the response to an impulse will agree well with the solution for the steady-state spectrum (Eq. \ref{eq:H0surf}). Dividing by the energy $E$, we find the probability distribution
\be \label{eq:spectrum}
P(\nu) & = &  -\frac{16\pi^2 R}{3k\phi E}  \sum_{nm} (-1)^n \gamma_{nm}^{-1} J_{nm}(\sigma).
\ee
Integrating over $\nu$ gives unity if the sum rule in Eq. \ref{eq:sumrule} is true.

@@

The distribution of the times it takes each photon to escape the sphere (hereafter called the ``wait time distribution'') is calculated from Eq. \ref{eq:dEdtdnu} by integrating over all frequencies. We obtain
\be
P(t)  & = & - \sqrt{\frac{3}{2}} \frac{16\pi^2 R \Delta^2 }{3kE}     \sum_{nm} (-1)^n  e^{s_{nm}t} \int d\sigma J_{nm}(\sigma) 
\nonumber \\ & = &  - \sum_{nm} P_{nm} s_{nm} e^{s_{nm}t},
\label{eq:waittime}
\ee
which is also normalized to unity.

>>>>

We now compare each of the solutions for surface flux with Monte Carlo simulations. In Fig. \ref{fig:sol_mc_residual_0}, the Monte Carlo is shown with solutions $H_{\rm d}$, $H_{\rm 0}$, and $H_{\rm 0 + bc} = H_{\rm 0} + H_{\rm bc}$ over-plotted for an optical depth of $\tau_0 = 10^7$ and photons emitted at line center $\rm x_s = 0$. The solution with the correct frequency-dependent boundary condition enforced, $H_{\rm 0 + bc}$, has considerably lower residuals to Monte Carlo than the other solutions, especially in the line wing. It is evident that the divergent solution $H_{\rm d}$ fails in the line wings. The $H_{\rm bc}$ term, when added to the solution $H_0$ derived by \citet{2006ApJ...649...14D}, adequately corrects for the apparent excess of flux in the peaks of the spectrum. This peak excess is also present in previous work that used the \citet{1990ApJ...350..216N} solution, such as \citet{2015MNRAS.449.4336S} where the discrepancy is particularly noticeable for line center optical depths of $\tau_0=10^5$ and $10^6$. The boundary term resolves the slight deficit of $H_{\rm0}$ in the line wings, further improving agreement with the numerical result. Note that the residuals to the $H_0$ solution are a close match to the $H_{\rm bc}$ term, since the Monte Carlo represents the ``true'' solution, $H$, and $H_{\rm bc} = H - H_0$.

@@

\begin{figure}
    \centering
    \includegraphics{xinit.pdf}
    \caption{$H_{\rm bc}$ solutions shown at three line center optical depths, $\tau_0=10^5, 10^7,$ and $10^9$, for $\sigma_s=0.0, \tau_0, 2\tau_0,$ etc. The axes are scaled to show that the size of the corrective factor agrees with the predicted $(a\tau_0)^{-1/3}$. scaling}
    \label{fig:xinit}
\end{figure}

Fig. \ref{fig:xinit} shows how the boundary term scales with both optical depth and photon source frequency. The photon initialization frequency is shifted in integer multiples of $\tau_0$. For clarity, only the positive side of the spectrum is shown. From Eq. \ref{eq:hbc_scaling}, the size of the correction should become smaller with larger optical depths, following a $(a\tau_0)^{-1/3}$ scaling. This factor has been scaled out of the figure such that solutions for different optical depths should show close agreement in scale on the figure if the relation holds, even while changing the source frequency of the photons. Indeed, the solutions converge as the optical depth becomes larger. The remaining discrepancy present is due to the width of the Doppler core with respect to the peak of the spectrum; at lower depths, this causes our assumptions about the functional form of the line profile to break down.

@@

Integrating $dE/dt d\nu$ from Eq. \ref{eq:dEdtdnu} over all times should give a distribution for the emitted frequencies. It is expected that this time-integrated spectrum of the response to an impulse will agree well with the solution for the $J=0$ steady-state spectrum (Eq. \ref{eq:H0surf}). Dividing by the energy $E$, we find the probability distribution
\be \label{eq:spectrum}
P(\nu) & = &  -\frac{16\pi^2 R}{3k\phi E}  \sum_{nm} (-1)^n \gamma_{nm}^{-1} J_{nm}(\sigma).
\ee
Integrating over $\nu$ gives unity if the sum rule in Eq. \ref{eq:sumrule} is true. In Fig. \ref{fig:steadystate}

\begin{figure}
    \centering
    \includegraphics{steadystate.pdf}
    \caption{Comparison of steady-state and time-integrated spectra for 20 spatial eigenmodes and 500 frequency eigenmodes. The accuracy of the red scatter points is limited by the number of frequency eigenmodes summed. The accuracy of the core behavior for the orange solid curve and red scatter points is limited by the number of spatial eigenmodes summed, which is why both solutions fail in the core.}
    \label{fig:my_label}
\end{figure}

@@

The distribution of the times it takes each photon to escape the sphere is calculated from Eq. \ref{eq:dEdtdnu} by integrating over all frequencies. We obtain
\be
P(t)  & = & - \sqrt{\frac{3}{2}} \frac{16\pi^2 R \Delta^2 }{3kE}     \sum_{nm} (-1)^n  e^{s_{nm}t} \int d\sigma J_{nm}(\sigma) 
\nonumber \\ & = &  - \sum_{nm} P_{nm} s_{nm} e^{s_{nm}t},
\label{eq:waittime}
\ee
which is also normalized to unity.

The eigenfunction expansion method is an effective characterization of the time-to-escape, yet not a fast enough computational method to use as a distribution for sampling. However, the behavior of the solution is well-understood since the late-time distribution is simply an exponential falloff in photon escape probability. The $e$-folding of the function is the lowest-order damping rate, $\gamma_{00}$. This is shown in Fig. \ref{fig:escape_time}. The orange line, representing the exponential falloff piece, matches well with the late-time behavior past a point $ct/R = ct_{0}/R = (a\tau_0)^{1/3}$. The early-time behavior can be modeled well with a gamma distribution, provided an adequate fit. For sampling, however, these fitting coefficients would need to be derived as functions of the input parameters $\tau_0$, $a$, and $x_s$.

\begin{figure}
    \centering
    \includegraphics{waittime.pdf}
    \caption{A wait time distribution of escaping photons, which is fit using the lowest order eigenfrequency as a decay timescale for the exponential falloff. The solid blue curve is a full solution using a sum over $n$ and $m$ modes. A gamma function can also be found that characterizes the early time behavior well, but underestimates photons escaping at late times.}
    \label{fig:escape_time}
\end{figure}
